### 2016-05-22 Paper/paper.tex
methods comments

\section{Introduction}
\citet{Sykulski2016} presents a multi-component spectral model for analysis of data transmitted by ocean surface drifters. 
Drifters are free-floating  instruments that transmit their location at regular time intervals,  creating a \textbf{\it{Langrangian time series}}, or a sequence of spatial locations over time. 
Figure ~\ref{fig: fig1} shows a sample 250-day drifter trajectory. 
Oceanographers use these time series to increase understanding of  ocean circulation patterns.
 Published in January 2016 in the \textbf{\it{Journal of the Royal Statistical Society Series C}}, this paper develops a model that can isolate a scientific quantity of interest,  and, in so doing, introduces several new statistical techniques.
 \begin{figure}[h!]
  \centering
    \includegraphics[width=.6\textwidth]{ReplicatedFigures/fig1.pdf}
        \caption{A 250-day sample drifter trajectory with two 50-day periods highlighted in green and blue. The time period highlighted in green is not affected by an eddy while the time period highlighted  in blue is.}
        	\label{fig: fig1}
\end{figure}

@@

	\subsection{Spectral Analysis}
	\label{sec: specAnalysis}
	\indent We briefly review spectral analysis as all the analysis in \citet{Sykulski2016} is done in the spectral domain.   
	More thorough coverage of this material can be found in  \citet{Percival1993}, for example. 
	From a statistical perspective, a time series, $z(t)$, real or complex-valued,  can be understood as a realization of a stochastic process over time, where  $t = \{1,2,...,n\}$ are the observed time points.  
	Time series are often modeled as the sums of periodic functions, which, using Euler's formula, can be represented compactly as Fourier series. 
	Moreover, data in the time domain can easily be related to data in the frequency domain and vice versa using the Fourier or inverse Fourier transformations 	\begin{align}
z(t) = \int_{-\infty}^{\infty} f_{z}(\omega)e^{i\omega t}d\omega && f_{z}(\omega) = \int_{-\infty}^{\infty} z(t) e^{-i \omega t }dt.
\end{align}
\citep{Percival1993}. 
To understand the distribution of frequencies, the \textbf{\it{power spectral density}},
\begin{equation}
S_{z}(\omega) = \underset{T \rightarrow \infty}{\lim} \mathbb{E} \left(\frac{1}{2T} \left| \int_{-T}^{T} z(t) e^{-i \omega t}dt \right|^{2} \right),
\end{equation}
is often employed. 
This quantity represents the variance (or power) associated with each frequency \citep{Percival1993}. 
Power spectral densities are useful for statistical modeling, since they can be related to the autocovariance sequence of a time series using  Fourier transformation.  
For real $z(t)$, we define the autocovariance as $s_{z}(\tau) = \mathbb{E}[z(t) z(t - \tau)] $ and for complex $z(t)$, we define $s_{z}(\tau) = \mathbb{E}[z(t) z^{*}(t - \tau)] $ where $z^{*}(t)$ represents the complex conjugate of z(t). Letting $\tau$ be the time lag, we have
\begin{align}
\label{eq: fourierPair}
S_{z}(\omega) = \int_{-\infty}^{\infty}s_{z}(\tau) e^{-\omega t}d \tau \Longleftrightarrow s_{z}(\tau) = \frac{1}{2\pi} \int_{-\infty}^{\infty}S_{z}(\omega) e^{i \omega t} d\omega 
\end{align}

@@

\subsubsection{Inertial oscillations}
We begin by discussing the model for the inertial oscillations. 
In a deterministic setting, inertial oscillations are often modeled with the following set of coupled differential equations
\begin{align}
\label{eq: deterOsc}
\frac{\partial u }{\partial t}  + f_{0} \nu &= F - cu \\ \nonumber
\frac{\partial v}{\partial t} - f_{0}u &= G - cv
\end{align}
where $u$ and $v$ again represent eastward and northward velocities, $f_{0}$ represents the inertial frequency in radians per unit time, and $F$ and $G$ are forces related to the wind  \citep{Pollard1970}. 
Suppressing the dependence on $t$ and using the complex representation of the velocity,  we show that these relationships can be equivalently expressed as follows
\begin{align}
\label{eq:diffEqDeriv}
\frac{dz}{dt} &= \frac{\partial u}{dt} + i\frac{\partial v(t)}{dt}\\ \nonumber
dz &= (F - c u- f_{0}v)dt + (iG - icv + if_{0}u)dt\\ \nonumber
&= fo(-v + iu)dt - c(u + iv)dt + (F + iG)dt\\ \nonumber
&= if_{0}(u + iv) - c(u + iv) + (F + iG)\\ \nonumber
&= (if_{0} - c)z dt + (F + iG)dt. 
\end{align}
\citet{Sykulski2016} further replace the wind forcing term, $(F + iG)dt$  with complex-valued Brownian increments \citep{Mandelbrot1968} to obtain a stochastic analogue for Equation ~\ref{eq: deterOsc},
 \begin{equation}
\label{eq: ouEq}
dz(t) = (i f_{0} -c) z(t) dt + A d Q(t). 
\end{equation}  
Equation ~\ref{eq: ouEq}  is known as the complex-valued Ornstein-Uhlenbeck (OU) stochastic differential equation. 
In general,  OU processes are considered continuous-time equivalents to the AR(1) process. Specifically,  the univariate OU process is the only mean-zero, stationary, Gaussian, and continuous Markov process, and, similarly the multivariate OU process is the only mean-zero $K$-variate Markov process that is stationary, continuous, Gaussian, and Markov \citep{Schach1971}.
 The complex-valued OU process is the complex representation of a bivariate OU process.  
 In the formulation of Equation ~\ref{eq: ouEq} , $c > 0$ is a dampening parameter that forces the process to always eventually return to its mean position, $z = 0$.  
 The final term of Equation ~\ref{eq: ouEq} is a positive constant, $A$, multiplied by $dQ(t)$.  The term $Q(t), t \geq 0$ is a standard complex Wiener process,  meaning $Q(t) = Q_{1}(t) + i Q_{2}(t)$, where $Q_{1}(t)$ and $Q_{2}(t)$ are standard Wiener processes.
 We remind the reader that Wiener processes are the stochastic models originally used to represent Brownian motion. 
  Such a process $\{Q(t, \omega): t \geq 0\}$ on the measure space $(\omega, Q, P)$ must satisfy
\begin{enumerate}
\item $Q(0, \omega) = 0$ almost everywhere
\item $\{Q(t, \omega): t \geq 0\}$ is distributed normally on $(\Omega,Q, P)$
\item $Q(t + \tau, \omega) - Q(t, \omega)$ has mean 0 and variance $|\tau|$ for all $t, \tau > 0$
\end{enumerate}
\citep{Hida1980}.

To identify and estimate shifts in the angular frequency, presumably caused by eddies, \citet{Sykulski2016} propose replacing the known inertial requency, $f_{0}$,  in Equation ~\ref{eq: ouEq} with  a free parameter, $\omega_{0}$, to be estimated by the data. 
If the estimated angular frequency differs from the inertial frequency, the angular frequency of the eddy can be calculated as $\omega_{eddy} = f_{0} - \omega_{0}$, since $\omega_{0}$ is a sum of the inertial and eddy frequencies.

For time series modeling, we are interested in the autocovariance of this process at different time lags. To derive this value given as Equation 4 of \citet{Sykulski2016}, we follow \citet{Arato1999} and consider the bivariate rather than complex-valued OU process.
This is possible since the autocovariance of a complex-valued random variable $z(t) = u(t) + i v(t)$ is defined as 
\begin{equation}
\label{eq: expec}
\mathbb{E}\{z(t) z^{*}(t + \tau)\} = \mathbb{E}[u(t)u(t + \tau)] + \mathbb{E}[(v(t) v(t + \tau)] + i\{\mathbb{E}[u(t) v(t + \tau)] + \mathbb{E}[u(t)v( t + \tau)]\}
\end{equation}
\citep{DeIaco2003} where $z^{*}(t + \tau)$ represents the complex conjugate of $z(t + \tau)$. The differential equations for components $u$ and $v$ can be expressed in matrix form as 
\begin{align}
Z(t) = \left( \begin{array}{c} u(t) \\ v(t) \end{array} \right) &= 
\left( \begin{array}{cc} -c & -\omega_{0} \\ \omega_{0} & -c \end{array} \right) \left( \begin{array}{cc} u(t) \\ v(t) \end{array} \right) + \left( \begin{array}{c} A d Q(t) \\ Ad Q(t) \end{array} \right). 
\end{align}
For a multivariate OU process with generator $B$, the covariance matrix for a lag of $\tau$ is $\exp \{B \tau \}C(0) $ where $s(0)$ represents the covariance at zero lag and $B$ is the generator matrix \citep{Schach1971}. 
Using the known variance of a univariate OU process and the independence of the components $u(0)$, we have
\begin{align}
s(0) = \left( \begin{array}{cc} A^{2}/2c & 0 \\ 0 & A^{2}/2c \end{array} \right).
\end{align}
Thus, 
\begin{align}
\label{eq: covDeriv}
\mathbb{E} \left[\left( \begin{array}{c} u(t + \tau) \\ v(t + \tau)  \end{array} \right) \left( \begin{array}{cc} u(t) & v(t) \end{array} \right)  \right] = e^{B \tau} s(0) = 
e^{-c \tau} \left( \begin{array}{cc} \cos \tau \omega_{0} & - \sin \tau \omega \\ \sin \tau \omega_{0} & \cos \tau \omega  \end{array} \right)  \left( \begin{array}{cc} A^{2}/2c & 0 \\ 0 & A^{2}/2c \end{array} \right)
\end{align}
where the second equality can be seen by expanding the infinite series representations for $\exp(-\lambda \tau)$, $\sin(\tau \omega)$, and $\cos( \tau \omega)$. 
Using the relevant components of Equation ~\ref{eq: covDeriv} in Equation ~\ref{eq: expec} and applying Euler's formula, we obtain the complex OU process autocovariance as defined in \citet{Sykulski2016}
\begin{align}
\label{eq: ouAC}
s^{(0)}(\tau) & = \frac{A^{2}}{2c} \exp(i \omega_{0}\tau) \exp(-c|\tau|).
\end{align}
 Using Equation ~\ref{eq: fourierPair}, the corresponding power spectral density can be found to be\begin{align}
\label{eq:ouPSD}
S^{(0)}(\omega) &= \int_{-\infty}^{\infty} s^{(0)}(\tau) \exp (-i \omega \tau) d \tau = \frac{A^{2}}{(\omega - \omega_{0}) + c^{2}}. 
\end{align}
The $(o)$ subscripts denotes that these equations are for the OU component of the full model.

\subsubsection{Turbulent Background}
The drifters' time series are also marked by the ocean's large-scale turbulence. 
The physics of ocean turbulence is summarized in \citet{Rhines1979}. 
Many models for ocean turbulence have been proposed, including several stochastic models \citep{Lacasce2008}. 
Within the oceanography literature, the stochastic models are designated  first-order \citep{Griffa1995, Falco2000}, second-order \citep{Sawford1991}, and third-order corresponding to  whether the velocity, acceleration, or hyperacceleration is modeled with a Markov process. 
However, calculation of the spectral slopes in observed data indicate that this class of models, referred to as \textbf{\it{integer order}},  does not always adequately reflect the ocean's behavior \citep{Rupolo1996, Sanderson1991}. Rather, \textbf{\it{fractional power models}}, such as fractional Brownian motion (FBM), would better reflect the observed behavior in some cases.

Towards this end, the authors propose using a Mat\'{e}rn covariance structure \citep{Gneiting2012} for the turbulent background. 
The Mat\'{e}rn model has gained prominence in the spatial statistics literature due to its flexibility.
 In this context, the Mat\'{e}rn model proves useful, since it allows for both integer and fractional powers. 
 The Mat\'{e}rn model is defined either by its autocovariance or power spectral density
\begin{align}
\label{eq:maternAC}
s^{(m)}(\tau) &= \frac{B^{2}}{2^{\alpha - 1/2}\pi^{1/2} \Gamma(\alpha) h^{2 \alpha - 1}}(h|\tau|)^{\alpha - 1}\kappa_{\alpha - 1/2}(h|\tau|)\\
\label{eq:maternPSD}
S^{(m)}(\omega) &= \frac{B^{2}}{(\omega^{2} + h^{2})^{\alpha}}
\end{align}
where $\tau$ again denotes the time lag, $\Gamma(\alpha)$ is the gamma function,  $\kappa_{\eta}$ is a modified Bessel function of the second kind of order $\eta$  and the subscript $(m)$ indicates that this is the Mat\'{e}rn part of the model \citep{Stein2012}. 
The parameters to be estimated are the amplitude, $B$, the dampening, $h > 0$, and a smoothness parameter $\alpha > 1/2$.
 The parameter $\alpha$ is of particular importance, since it controls the smoothness, or degree of differentiability, of the process \citep{Fuentes2010}.

The authors emphasize that the turbulent background models previously proposed are special cases of the Mat\'{e}rn model. 
This means that these models can still be obtained under the current model if model fitting leads to particular parameter estimate values. 
They also highlight that that Mat\'{e}rn model is more appropriate than the other obvious choice of a fractional power model, FBM, since FBM does not have a bound on the velocities and tends to drift quite far from its mean.
 Figure ~\ref{fig: fmbMat} illustrates these differences by comparing the paths of a simulated particle obtained with a Mat\'{e}rn  covariance versus a FBM covariance.

\begin{figure}[h!]
  \centering
    \includegraphics[width=.7\textwidth]{ReplicatedFigures/fig3.pdf}
        \caption{Simulated velocities obtained using FBM (blue) and Mat\'{e}rn (red)  processes. Top Left: The eastward component the velocities. Top Right: The eastward and northward components of the velocities. Bottom: Periodograms of both processes on the decibel scale. }
        	\label{fig: fmbMat}
\end{figure}

\subsubsection{Overall Model}
Combining the inertial oscillation and turbulent background parts of the model is straightforward.
 To find both the autocovariance and power spectral densities, we can simply add the corresponding components of the inertial oscillation and turbulent background autocovariances. 
 Using Equations ~\ref{eq: ouAC} and ~\ref{eq:maternAC} and Equations ~\ref{eq:ouPSD} and ~\ref{eq:maternPSD}, we obtain
\begin{align}
S(\omega; \boldsymbol{\theta}) &= \frac{A^{2}}{(\omega - \omega_{0}) + c^{2}} + \frac{B^{2}}{(\omega^{2} + h^{2})^{\alpha}}\\
\label{eq: fullSpec}
s_{\tau}(\boldsymbol{\theta}) &= \frac{A^{2}}{2c} \exp(i \omega_{0}\tau) \exp(-c|\tau|) +  \frac{B^{2}}{2^{\alpha - 1/2}\pi^{1/2} \Gamma(\alpha) h^{2 \alpha - 1}}(h|\tau|)^{\alpha - 1}\kappa_{\alpha - 1/2}(h|\tau|).
\end{align}

@@

\subsection{Fitting the Model}
\subsubsection{Blurred Whittle likelihood}
\par A recent technique introduced in  \citet{Sykulski2013}  is used to fit the model.
 Referred to as \textbf{\it{blurred Whittle likelihood}}, this new likelihood builds on the well-known Whittle likelihood, a frequency-based approximation  to the time-domain likelihood that reduces computational cost. 
  Whittle likelihood is designed to approximate the likelihood for a univariate Gaussian process, $Z$,  in the time domain that has the likelihood
\begin{align*}
l_{t}(\theta) = - \frac{1}{2} log |C_{Z}(\theta) | - \frac{1}{2} Z^{T} C_{Z}^{-1}(\theta)Z
\end{align*}
where $\theta$ are the model parameters for a particular choice of covariance structure  and $t$ denotes the time domain  \citep{Sykulski2013}.

 It is known that  the periodogram, given in Equation ~\ref{eq: perio}, satisfies
\begin{equation}
|J_{Z}(\omega)| \overset{d}{\rightarrow} S_{Z}(\omega) \chi^{2}_{2}/2
\end{equation}
under certain regularity conditions \citep{Contreras2006}. 
\citet{Whittle1953} combined this result with Fourier approximations to the true covariance structure and features of Toeplitz matrices to obtain the approximate likelihood
\begin{align}
l_{\omega}(\theta) &= - \sum_{\omega \in \Omega} \left[ J_{Z}^{H}(\omega) \tilde{S}_{Z}^{-1}(\omega; \theta) J_{Z}(\omega) + \log \{S(\omega; \theta) \} \right]\\
\label{eq: bwl}
&= -\sum_{\omega \in \Omega} \left[ \frac{\hat{S}_{Z}(\omega)}{S(\omega;\theta)}  + \log  \{ S(\omega; \theta) \}\right]
\end{align}
 where the subscript $w$ indicates that this is a Whittle likelihood and $H$ denotes a Hermitian matrix. 
  This likelihood can be evaluated in $\mathcal{O}(N log(N))$ where $N$ is the length of the time series which is a notable increase in speed compared to the $\mathcal{O}(N^{3})$  required for time-domain likelihood  \citep{Sykulski2016}. 
  Nevertheless, this likelihood relies on the periodogram and so is affected by aliasing and leakage as discussed in Section ~\ref{sec: specAnalysis}.

>>>>

\section{Introduction}
\citet{Sykulski2016} present\vmdel{s} a multi-component spectral model for analysis of data transmitted by ocean surface drifters. 
Drifters are free-floating  instruments that transmit their location at regular time intervals,  creating a \textbf{\it{Langrangian time series}}, or a sequence of spatial locations over time. 
Figure~\ref{fig: fig1} shows a sample 250-day drifter trajectory. 
Oceanographers use these time series to increase understanding of  ocean circulation patterns.
Published in January 2016 in the \textbf{\it{Journal of the Royal Statistical Society Series C}}, this paper develops a model that can isolate a scientific quantity of interest,  and, in so doing, introduces several new statistical techniques. \vmcomment{The last sentence is very low on information content.}
 \begin{figure}[h!]
  \centering
    \includegraphics[width=.6\textwidth]{ReplicatedFigures/fig1.pdf}
        \caption{A 250-day sample drifter trajectory with two 50-day periods highlighted in green and blue. The time period highlighted in green is not affected by an eddy while the time period highlighted  in blue is.}
        	\label{fig: fig1}
\end{figure}

@@

	\subsection{Spectral Analysis}
	\label{sec: specAnalysis}
	\indent We briefly review spectral analysis as all the analysis in \citet{Sykulski2016} is done in the spectral domain.   
	More thorough coverage of this material can be found in  \citep{Percival1993}, for example. 
	From a statistical perspective, a time series, $z(t)$, real or complex-valued,  can be understood as a realization of a stochastic process over time, where  $t = \{1,2,...,n\}$ are the observed time points.  
	Time series are often modeled as the sums of periodic functions, which, using Euler's formula, can be represented compactly as Fourier series. 
	Moreover, data in the time domain can easily be related to data in the frequency domain and vice versa using the Fourier or inverse Fourier transformations\vmadd{:} 	
	\begin{align}
z(t) = \int_{-\infty}^{\infty} f_{z}(\omega)e^{i\omega t}d\omega\vmadd{,} && f_{z}(\omega) = \int_{-\infty}^{\infty} z(t) e^{-i \omega t }dt.
\end{align}
\citep{Percival1993}. 
To understand the distribution of frequencies, the \textbf{\it{power spectral density}},
\begin{equation}
S_{z}(\omega) = \underset{T \rightarrow \infty}{\lim} \mathbb{E} \left(\frac{1}{2T} \left| \int_{-T}^{T} z(t) e^{-i \omega t}dt \right|^{2} \right),
\end{equation}
is often employed. 
This quantity represents the variance (or power) associated with each frequency \citep{Percival1993}. 
Power spectral densities are useful for statistical modeling, since they can be related to the autocovariance sequence of a time series using  Fourier transformation.  
For real $z(t)$, we define the autocovariance as $s_{z}(\tau) = \mathbb{E}[z(t) z(t - \tau)] $ and for complex $z(t)$, we define $s_{z}(\tau) = \mathbb{E}[z(t) z^{*}(t - \tau)] $ where $z^{*}(t)$ represents the complex conjugate of z(t). Letting $\tau$ be the time lag, we have
\begin{align}
\label{eq: fourierPair}
S_{z}(\omega) = \int_{-\infty}^{\infty}s_{z}(\tau) e^{-\omega t}d \tau \Longleftrightarrow s_{z}(\tau) = \frac{1}{2\pi} \int_{-\infty}^{\infty}S_{z}(\omega) e^{i \omega t} d\omega 
\end{align}

@@

\subsubsection{Inertial oscillations}
We begin by discussing the model for the inertial oscillations. 
In a deterministic setting, inertial oscillations are often modeled with the following set of coupled differential equations
\begin{align}
\label{eq: deterOsc}
\frac{\partial u }{\partial t}  + f_{0} \nu &= F - cu\vmadd{,} \\ \nonumber
\frac{\partial v}{\partial t} - f_{0}u &= G - cv\vmadd{,}
\end{align}
where $u$ and $v$ again represent eastward and northward velocities, $f_{0}$ represents the inertial frequency in radians per unit time, and $F$ and $G$ are forces related to the wind  \citep{Pollard1970}. 
Suppressing the dependence on $t$ and using the complex representation of the velocity,  we show that these relationships can be equivalently expressed as follows
\begin{align}
\label{eq:diffEqDeriv}
\frac{dz}{dt} &= \frac{\partial u}{dt} + i\frac{\partial v(t)}{dt}\\ \nonumber
dz &= (F - c u- f_{0}v)dt + (iG - icv + if_{0}u)dt\\ \nonumber
&= fo(-v + iu)dt - c(u + iv)dt + (F + iG)dt\\ \nonumber
&= if_{0}(u + iv) - c(u + iv) + (F + iG)\\ \nonumber
&= (if_{0} - c)z dt + (F + iG)dt. 
\end{align}
\citet{Sykulski2016} further replace the wind forcing term, $(F + iG)dt$  with complex-valued Brownian increments \citep{Mandelbrot1968} to obtain a stochastic analogue for Equation ~\ref{eq: deterOsc},
 \begin{equation}
\label{eq: ouEq}
dz(t) = (i f_{0} -c) z(t) dt + A d Q(t). 
\end{equation}  
Equation ~\ref{eq: ouEq}  is known as the complex-valued Ornstein-Uhlenbeck (OU) stochastic differential equation. 
In general,  OU processes are considered continuous-time equivalents to the AR(1) process. Specifically,  the univariate OU process is the only mean-zero, stationary, Gaussian, and continuous Markov process, and, similarly the multivariate OU process is the only mean-zero $K$-variate Markov process that is stationary, continuous, Gaussian, and Markov \citep{Schach1971}.
 The complex-valued OU process is the complex representation of a bivariate OU process.  
 In the formulation of Equation~\ref{eq: ouEq} , $c > 0$ is a dampening parameter that forces the process to always eventually return to its mean position, $z = 0$.  
 The final term of Equation~\ref{eq: ouEq} is a positive constant, $A$, multiplied by $dQ(t)$.  The term $Q(t), t \geq 0$ is a standard complex Wiener process,  meaning $Q(t) = Q_{1}(t) + i Q_{2}(t)$, where $Q_{1}(t)$ and $Q_{2}(t)$ are standard Wiener processes.
 We remind the reader that Wiener processes are the stochastic models originally used to represent Brownian motion. 
  Such a process $\{Q(t, \omega): t \geq 0\}$ on the measure space $(\omega, Q, P)$ must satisfy
\begin{enumerate}
\item $Q(0, \omega) = 0$ almost everywhere
\item $\{Q(t, \omega): t \geq 0\}$ is distributed normally on $(\Omega,Q, P)$
\item $Q(t + \tau, \omega) - Q(t, \omega)$ has mean 0 and variance $|\tau|$ for all $t, \tau > 0$
\end{enumerate}
\citep{Hida1980}.

To identify and estimate shifts in the angular frequency, presumably caused by eddies, \citet{Sykulski2016} propose replacing the known inertial requency, $f_{0}$,  in Equation~\ref{eq: ouEq} with  a free parameter, $\omega_{0}$, to be estimated by the data. 
If the estimated angular frequency differs from the inertial frequency, the angular frequency of the eddy can be calculated as $\omega_{eddy} = f_{0} - \omega_{0}$, since $\omega_{0}$ is a sum of the inertial and eddy frequencies.

For time series modeling, we are interested in the autocovariance of this process at different time lags. To derive this value given as Equation 4 of \citet{Sykulski2016}, we follow \citet{Arato1999} and consider the bivariate rather than complex-valued OU process.
This is possible since the autocovariance of a complex-valued random variable $z(t) = u(t) + i v(t)$ is defined as 
\begin{equation}
\label{eq: expec}
\mathbb{E}\{z(t) z^{*}(t + \tau)\} = \mathbb{E}[u(t)u(t + \tau)] + \mathbb{E}[(v(t) v(t + \tau)] + i\{\mathbb{E}[u(t) v(t + \tau)] + \mathbb{E}[u(t)v( t + \tau)]\}
\end{equation}
\citep{DeIaco2003} where $z^{*}(t + \tau)$ represents the complex conjugate of $z(t + \tau)$. The differential equations for components $u$ and $v$ can be expressed in matrix form as 
\begin{align}
Z(t) = \left( \begin{array}{c} u(t) \\ v(t) \end{array} \right) &= 
\left( \begin{array}{cc} -c & -\omega_{0} \\ \omega_{0} & -c \end{array} \right) \left( \begin{array}{cc} u(t) \\ v(t) \end{array} \right) + \left( \begin{array}{c} A d Q(t) \\ Ad Q(t) \end{array} \right). 
\end{align}
For a multivariate OU process with generator $B$, the covariance matrix for a lag of $\tau$ is $\exp \{B \tau \}C(0)$\vmadd{,} where $s(0)$ represents the covariance at zero lag \vmdel{and $B$ is the generator matrix} \citep{Schach1971}. \vmcomment{Generator is defined neither in general nor in your particular case.}
Using the known variance of a univariate OU process and the independence of the components $u(0)$, we have
\begin{align}
s(0) = \left( \begin{array}{cc} A^{2}/2c & 0 \\ 0 & A^{2}/2c \end{array} \right).
\end{align}
Thus, 
\begin{align}
\label{eq: covDeriv}
\mathbb{E} \left[\left( \begin{array}{c} u(t + \tau) \\ v(t + \tau)  \end{array} \right) \left( \begin{array}{cc} u(t) & v(t) \end{array} \right)  \right] = e^{B \tau} s(0) = 
e^{-c \tau} \left( \begin{array}{cc} \cos \tau \omega_{0} & - \sin \tau \omega \\ \sin \tau \omega_{0} & \cos \tau \omega  \end{array} \right)  \left( \begin{array}{cc} A^{2}/2c & 0 \\ 0 & A^{2}/2c \end{array} \right)\vmadd{,}
\end{align}
where the second equality can be seen by expanding the infinite series representations for $\exp(-\lambda \tau)$, $\sin(\tau \omega)$, and $\cos( \tau \omega)$. 
Using the relevant components of Equation ~\ref{eq: covDeriv} in Equation ~\ref{eq: expec} and applying Euler's formula, we obtain the complex OU process autocovariance as defined in \citet{Sykulski2016}
\begin{align}
\label{eq: ouAC}
s^{(0)}(\tau) & = \frac{A^{2}}{2c} \exp(i \omega_{0}\tau) \exp(-c|\tau|).
\end{align}
 Using Equation ~\ref{eq: fourierPair}, the corresponding power spectral density can be found to be\begin{align}
\label{eq:ouPSD}
S^{(0)}(\omega) &= \int_{-\infty}^{\infty} s^{(0)}(\tau) \exp (-i \omega \tau) d \tau = \frac{A^{2}}{(\omega - \omega_{0}) + c^{2}}. 
\end{align}
The $(o)$ subscripts denotes that these equations are for the OU component of the full model.

\subsubsection{Turbulent Background}
The drifters' time series are also marked by the ocean's large-scale turbulence. 
The physics of ocean turbulence is summarized \vmdel{in} \vmadd{by} \citet{Rhines1979}. 
Many models for ocean turbulence have been proposed, including several stochastic models \citep{Lacasce2008}. 
Within the oceanography literature, the stochastic models are designated  first-order \citep{Griffa1995, Falco2000}, second-order \citep{Sawford1991}, and third-order corresponding to  whether the velocity, acceleration, or hyperacceleration is modeled with a Markov process. 
However, calculation of the spectral slopes in observed data indicate that this class of models, referred to as \textbf{\it{integer order}},  does not always adequately reflect the ocean's behavior \citep{Rupolo1996, Sanderson1991}. Rather, \textbf{\it{fractional power models}}, such as fractional Brownian motion (FBM), would better reflect the observed behavior in some cases.

Towards this end, the authors propose using a Mat\'{e}rn covariance structure \citep{Gneiting2012} for the turbulent background. 
The Mat\'{e}rn model has gained prominence in the spatial statistics literature due to its flexibility.
 In this context, the Mat\'{e}rn model proves useful, since it allows for both integer and fractional powers. 
 The Mat\'{e}rn model is defined either by its autocovariance \vmdel{or power spectral density}
\begin{align}
\label{eq:maternAC}
s^{(m)}(\tau) = \frac{B^{2}}{2^{\alpha - 1/2}\pi^{1/2} \Gamma(\alpha) h^{2 \alpha - 1}}(h|\tau|)^{\alpha - 1}\kappa_{\alpha - 1/2}(h|\tau|)\vmadd{,}
\end{align}
\vmadd{or power spectral density,}
\begin{align}
\label{eq:maternPSD}
S^{(m)}(\omega) = \frac{B^{2}}{(\omega^{2} + h^{2})^{\alpha}}\vmadd{,}
\end{align}
where $\tau$ again denotes the time lag, $\Gamma(\alpha)$ is the gamma function,  $\kappa_{\eta}$ is a modified Bessel function of the second kind of order $\eta$  and the s\vmdel{ub}\vmadd{uper}script $(m)$ indicates that this is the Mat\'{e}rn part of the model \citep{Stein2012}. 
The parameters to be estimated are the amplitude, $B$, the dampening, $h > 0$, and a smoothness parameter $\alpha > 1/2$.
 The parameter $\alpha$ is of particular importance, since it controls the smoothness, or degree of differentiability, of the process \citep{Fuentes2010}.

The authors emphasize that the turbulent background models previously proposed are special cases of the Mat\'{e}rn model. 
This means that these models can still be obtained under the current model if model fitting leads to particular parameter estimate values. 
They also highlight that that Mat\'{e}rn model is more appropriate than the other obvious choice of a fractional power model, FBM, since FBM does not have a bound on the velocities and tends to drift quite far from its mean.
 Figure~\ref{fig: fmbMat} illustrates these differences by comparing the paths of a simulated particle obtained with a Mat\'{e}rn  covariance versus a FBM covariance.

\begin{figure}[h!]
  \centering
    \includegraphics[width=.7\textwidth]{ReplicatedFigures/fig3.pdf}
        \caption{Simulated velocities obtained using FBM (blue) and Mat\'{e}rn (red)  processes. Top Left: The eastward component the velocities. Top Right: The eastward and northward components of the velocities. Bottom: Periodograms of both processes on the decibel scale. \vmcomment{I would include legends or direct labeling for all your figures.}}
        	\label{fig: fmbMat}
\end{figure}

\subsubsection{Overall Model}
Combining the inertial oscillation and turbulent background parts of the model is straightforward.
 To find both the autocovariance and power spectral densities, we can simply add the corresponding components of the inertial oscillation and turbulent background autocovariances. 
 Using Equations ~\ref{eq: ouAC} and ~\ref{eq:maternAC} and Equations ~\ref{eq:ouPSD} and ~\ref{eq:maternPSD}, we obtain
\begin{align}
S(\omega; \boldsymbol{\theta}) &= \frac{A^{2}}{(\omega - \omega_{0}) + c^{2}} + \frac{B^{2}}{(\omega^{2} + h^{2})^{\alpha}}\vmadd{,}\\
\label{eq: fullSpec}
s_{\tau}(\boldsymbol{\theta}) &= \frac{A^{2}}{2c} \exp(i \omega_{0}\tau) \exp(-c|\tau|) +  \frac{B^{2}}{2^{\alpha - 1/2}\pi^{1/2} \Gamma(\alpha) h^{2 \alpha - 1}}(h|\tau|)^{\alpha - 1}\kappa_{\alpha - 1/2}(h|\tau|).
\end{align}

@@

\subsection{Fitting the Model}
\subsubsection{Blurred Whittle likelihood}
\par A recent technique introduced by \citet{Sykulski2013}  is used to fit the model.
 Referred to as \textbf{\it{blurred Whittle likelihood}}, this new likelihood builds on the well-known Whittle likelihood, a frequency-based approximation  to the time-domain likelihood that reduces computational cost. 
  Whittle likelihood is designed to approximate the likelihood for a univariate Gaussian process, $Z$,  in the time domain that has the likelihood
\begin{align*}
l_{t}(\theta) = - \frac{1}{2} log |C_{Z}(\theta) | - \frac{1}{2} Z^{T} C_{Z}^{-1}(\theta)Z\vmadd{,}
\end{align*}
where $\theta$ are the model parameters for a particular choice of covariance structure  and $t$ denotes the time domain  \citep{Sykulski2013}. \vmcomment{I should have mentioned this earlier in the test that all vectors and matrices need to be bolded.}

 It is known that  the periodogram, given in Equation ~\ref{eq: perio}, satisfies
\begin{equation}
|J_{Z}(\omega)| \overset{d}{\rightarrow} S_{Z}(\omega) \chi^{2}_{2}/2
\end{equation}
under certain regularity conditions \citep{Contreras2006}. 
\citet{Whittle1953} combined this result with Fourier approximations to the true covariance structure and features of Toeplitz matrices to obtain the approximate likelihood
\begin{align}
l_{\omega}(\theta) &= - \sum_{\omega \in \Omega} \left[ J_{Z}^{H}(\omega) \tilde{S}_{Z}^{-1}(\omega; \theta) J_{Z}(\omega) + \log \{S(\omega; \theta) \} \right]\\
\label{eq: bwl}
&= -\sum_{\omega \in \Omega} \left[ \frac{\hat{S}_{Z}(\omega)}{S(\omega;\theta)}  + \log  \{ S(\omega; \theta) \}\right]\vmadd{,}
\end{align}
where the subscript $w$ indicates that this is a Whittle likelihood and $H$ denotes a Hermitian matrix. 
  This likelihood can be evaluated in $\mathcal{O}(N log(N))$ where $N$ is the length of the time series which is a notable increase in speed compared to the $\mathcal{O}(N^{3})$  required for time-domain likelihood  \citep{Sykulski2016}. 
  Nevertheless, this likelihood relies on the periodogram and so is affected by aliasing and leakage as discussed in Section~\ref{sec: specAnalysis}.

To improve on these problems, \citet{Sykulski2013} propose replacing the theoretical spectrum, $S_{Z}(\omega; \theta)$\vmadd{, in Equation ~\ref{eq: bwl}} with the expected value of the spectrum, \vmdel{$\overline{S}_{Z}(\omega, \theta)$}
\begin{equation}
\overline{S}(\omega; \theta) = \Delta \sum_{\tau = - (N - 1)}^{N-1} \left(1 - \frac{|\tau|}{N} \right)s_{\tau}(\theta) exp ( - i \omega \tau \Delta)\vmadd{.}
\end{equation} 
This form of the expected spectrum can be shown to be equivalent to convolving the theoretical spectrum with the Fe\'{j}er kernel 
\begin{equation}
\mathcal{F}(\cdot) = \frac{\Delta}{2\pi N} \frac{sin^{2}(N \omega \Delta/2)}{sin^{2}(\omega \Delta /2)} 
\end{equation} \citep{Sykulski2013}. 
This improves the estimators obtained through maximizing the likelihood, because the expected value of the periodogram $\mathbb{E}[|J_{Z}(\omega)^{2}]$ is blurred in the same way as the observed periodogram. Thereby, effects due to sampling at discrete intervals are avoided. 
This leads to the final likelihood
\begin{align}
l_{b}(\theta) = - \sum_{\omega \in \Omega} \left[\frac{\hat{S}_{Z}(\omega)}{\overline{S}(\omega; \theta)} + \log \{ \overline{S} (\omega; \theta) \}\right]\vmadd{,}
\end{align}
which is maximized by standard \vmadd{off-the-shelf nonlinear} optimization techniques. 
\vmcomment{State the exact method(s) that the paper and you used.}
